PyQBench: a Python library for benchmarking gate-based quantum computers\\ Supplemental material

### 2023-10-05 txt/manuscript_short.tex
Move section 2 to the appendix + conciseness pass

Noisy Intermediate-Scale Quantum (NISQ)~\cite{preskill} devices are storming the market, with a wide
selection of devices based on different architectures and accompanying software solutions. Among
hardware providers offering public access to their gate--based devices, one could mention Rigetti
\cite{rigetti}, IBM \cite{ibmq}, Oxford Quantum Group \cite{oxforf}, IonQ \cite{ionq} or Xanadu
\cite{xanadu}. Other vendors offer devices operating in different paradigms. Notably, one could
mention D-Wave \cite{dwave} and their quantum annealers, or QuEra devices \cite{quera} based on
neural atoms. Most vendors provide their own software stack and application programming interface
for accessing their devices. To name a few, Rigetti's computers are available through their Forest
SDK \cite{sdk} and PyQuil library \cite{pyquil} and IBM Q \cite{ibmq} computers can be accessed
through Qiskit \cite{qiskit} or IBM Quantum Experience web interface \cite{ibmqplatform}. Some cloud
services, like Amazon Braket \cite{amazon}, offer access to several quantum devices under a unified
API. On top of that, several libraries and frameworks can integrate with multiple hardware vendors.
Examples of such frameworks include IBM Q's Qiskit, Zapata Computing's Orquestra \cite{zapata}, XACC \cite{xacc} and NVIDIA's CUDA Quantum \cite{cudaquantum}.

It is well known that NISQ devices have their limitations \cite{preskillnew}. The question is to
what extent those devices can perform meaningful computations? To answer this question, one has to
devise a methodology for benchmarking them. For gate--based computers, on which this paper focuses,
there already exist several approaches. One could mention randomized benchmarking
\cite{liu2022sampling, knill2008randomized, wallman2014randomized, helsen2022general,
cornelissen2021scalable}, benchmarks based on the quantum volume \cite{cross2019validating,
moll2018quantum, pelofske2022volume}.

In this paper, we introduce a different approach to benchmarking gate--based devices with a simple
operational interpretation. In our method, we test how well the given device is at guessing which of
the two known von Neumann measurements were performed during the experiment. We implemented our
approach in an open-source Python library called PyQBench. The library supports any device available
through the Qiskit library, and thus can be used with providers such as IBM Q or Amazon Braket.
Along with the library, the PyQBench package contains a command line tool for running most common
benchmarking scenarios.

\section{Existing benchmarking methodologies and software}
Unsurprisingly, PyQBench is not the only software package for benchmarking gate--based devices.
While we believe that our approach has significant benefits over other benchmarking techniques, for
completeness, in this section we discuss some of the currently available similar software.

Probably the simplest benchmarking method one could devise is simply running known algorithms and
comparing outputs with the expected ones. Analyzing the frequency of the correct outputs, or the
deviation between actual and expected outputs distribution provides then a metric of the performance
of a given device. Libraries such as Munich Quantum Toolkit (MQT) \cite{mqt2022, mqt-bench} or
SupermarQ \cite{supermarq, supermarkq-github} contain benchmarks leveraging multiple algorithms,
such as Shor's algorithm or Grover's algorithm.
% or  implementing e.g. GHZ state \cite{ghz}.
Despite being intuitive and easily interpretable, such benchmarks may have some problems. Most
importantly, they assess the usefulness of a quantum device only for a very particular algorithm,
and it might be hard to extrapolate their results to other algorithms and applications. For
instance, the inability of a device to consistently find factorizations using Shor's algorithms does
not tell anything about its usefulness in Variational Quantum Algorithm's.

Another possible approach to benchmarking quantum computers is randomized benchmarking. In this
approach, one samples circuits to be run from some predefined set of gates (e.g. from the Clifford
group) and tests how much the output distribution obtained from the device running these circuits
differs from the ideal one. It is also common to concatenate randomly chosen circuits with their
inverses (which should yield the identity circuit) and run those concatenated circuits on the
device. Libraries implementing this approach include Qiskit \cite{qiskit-randomized} or PyQuil
\cite{forest-benchmarking}. Another, equally popular, benchmarking method is
quantum tomography \cite{bantysh2021quantum}. Implementation of quantum tomography for
benchmarking NISQ devices can be found in \cite{tomography}.

In \cite{patel2020experimental}, the authors evaluated several IBM-Q machines using seven benchmarks
taking into account the errors and execution time.

QASMBench \cite{li2023qasmbench} is one of the first benchmark suites aiming at evaluating NISQ
devices using quantum applications from a broad range of domains, mainly using an approach based on
fidelity estimation. The benchmark presented in the paper compares the fidelity of execution among
the IBM-Q machines, the IonQ QPU and the Rigetti Aspen M-1 system.

Another quantity used for benchmarking NISQ devices is quantum volume. The quantum volume
characterizes the capacity of a device for solving computational problems. It takes into account
multiple factors like the number of qubits, connectivity and measurement errors. The Qiskit library
allows one to measure the quantum volume of a device by using its \texttt{qiskit.ignis.verification.quantum\_volume}. Other implementations of Quantum Volume can be found as well, see e.g.
\cite{volume-in-practice}.

Finally, we should mention cross-entropy
benchmarking~\cite{boixo2018characterizing}, which was utlilized in validation
of the Sycamore-53 QPU supremacy experiments~\cite{arute2019quantum}. In this
approach, the quality of an algorithm implemented on the QPU is measured by
calculating the cross entropy of bit-strings actually sampled from the QPU,
compared to ideal bitstrings.

@@

In this section, we describe how the benchmarking process in PyQBench works. We start by
discussing necessary mathematical preliminaries. Then, we present the general form of the
discrimination scheme used in PyQBench and practical considerations on how to implement it taking
into account the limitations of the current NISQ devices. We encourage the readers interested
in a more in--depth discussion of the mathematical foundations behind our discrimination scheme to read
Section \ref{app:preliminaries} in the supplemental materials.

@@

A von Neumann measurement $\PP$ is a collection of rank--one projectors
$\{\proj{u_0}, \ldots, \proj{u_{d-1}}\}$, called effects, that sum up to the identity operator, i.e.
$ \, \, \sum_{i=0}^{d-1} \proj{u_i} = \1$. If $U$ is a unitary matrix of size $d$, one can construct
a von Neumann measurement $\PP_{U}$ by taking projectors onto its columns. In this case we say that
$\PP_{U}$ is described by the matrix $U$.

@@

Without loss of generality\footnote{Explaining why we can consider only discrimination scheme
between $\PP_{\Id}$ and $\PP_{U}$ is beyond the scope of this paper. See
\cite{puchala2018strategies} for a in depth explanation.}, we consider discrimination task between
single qubit measurements $\PP_\Id$, performed in the computational Z-basis, and an alternative
measurement $\PP_U$ performed in the basis $U$. 
The discrimination scheme presented in Fig.~\ref{fig:theoretical_scheme} requires an
auxiliary qubit. First, the joint system is prepared in some state $\ket{\psi_0}$. Then, one of the
measurements, either $\PP_U$ or $\PP_\1$, is performed on the first part of the system. Based on its
outcome $i$, we choose another binary measurement $\mathcal{P}_{V_i}$ and perform it on the second qubit,
obtaining the outcome $j$. Finally, if $j=0$, we say that the performed measurement is
$\mathcal{P}_U$, otherwise we say that it was $\mathcal{P}_\Id$.

Note, however, that the discrimination scheme described above can work regardless of dimensionality 
of the unitary $U$. The main difference is the dimension of the auxiliary system, which in general
can be larger than two. This dimension depends on the Schmidt rank of the optimal input state. However, 
for most discrimination schemes, the  Schmidt rank equals at most two, and hence the auxiliary system
is also a qubit, see \cite{puchala2018strategies} for details. Note that the final measurement $\PP_{V_i}$ 
is always binary, independently of the dimension of auxiliary system.

@@

Current NISQ devices are unable to perform conditional measurements, which is the biggest obstacle
to implementing our scheme on real hardware. However, we circumvent this problem by slightly
adjusting our scheme so that it only uses components available on current devices. For this purpose,
we use two possible options: using a postselection or a direct sum $V_0^\dagger\oplus V_1^\dagger$.

@@

	The first idea uses a postselection scheme. In the original scheme, we measure the first qubit
	and only then determine which measurement should be performed on the second one. Instead of doing
	this choice, we can run two circuits, one with $\PP_{V_0}$ and one with $\PP_{V_1}$ and measure both
	qubits. We then discard the results of the circuit for which label $i$ does not match measurement
	label $k$. Hence, the circuit for postselection looks as depicted in Fig.
	\ref{fig:postselection}.

@@

	One can see why such a circuit is equivalent to the original discrimination scheme.
	If we rewrite the block-diagonal matrix $V_0^\dagger \oplus V_1^\dagger$ as follows:
	\begin{equation}
		\label{eq:directsum}
		V_0^\dagger \oplus V_1^\dagger = \proj{0}\otimes V_0^\dagger + \proj{1} \otimes V_1^\dagger,
	\end{equation}
	we can see that the direct sum in Eq. \eqref{eq:directsum} commutes with the measurement on the
	first qubit. Thanks to this, we can switch the order of operations to obtain the circuit from
	Fig. \ref{fig:directsum}. Now, depending on the outcome $i$, one of the summands in
	Eq. \eqref{eq:directsum} vanishes, and we end up performing exactly the same operations as in
	the original scheme.

@@

Compared to these approaches, our approach allows for a very simple operational
interpretation of the scheme and its results. This is especially useful for
newcomers to the field, who may be put off by more complicated approaches.
Another benefit is, especially for advanced users, the ability to control
resources utilized during the benchmarking. We can consider here resources such
as entanglement or coherence. Finally, the figure of merit we wish to calculate and
to which we compare the results is fairly simple to obtain. The main downside of
our approach is the exponential number of circuits we need to consider.

@@

\begin{enumerate}
	\item Assembling circuits for both postselection and direct sum--based discrimination schemes.
	\item Executing the whole benchmarking scenario on specified backend (either real hardware or
	software simulator).
	\item Interpreting the obtained outputs in terms of discrimination probabilities.
\end{enumerate}
Note that the execution of circuits by PyQBench is optional. Instead, the user might want to opt in
for fine-grained control over the execution of the circuits. For instance, suppose the user wants to
simulate the discrimination experiment on a noisy simulator. In such a case, they can define the
necessary components and assemble the circuits using PyQBench. The circuits can then be altered,
e.g. to add noise to particular gates, and then run using any Qiskit backend by the user. Finally,
PyQBench can be used to interpret the measurements to obtain discrimination probability.

The PyQBench library also contains a readily available implementation of all necessary components
needed to run discrimination experiments for parametrized Fourier family of measurements (see
Section \ref{app:optimal-probability} in supplemental material). However, if one only wishes to use
this particular family of measurements in their benchmarks, then using PyQBench as a command line tool
might be more straightforward. PyQBench's command line interface allows running the benchmarking process
without writing Python code. The configuration of CLI is done by YAML \cite{yaml} files describing the
benchmark to be performed and the description of the backend on which the benchmark should be run.
The same benchmark can be used with different backends and vice versa.

@@

\subsubsection{Overview of the software structure}
As already described, PyQBench can be used both as a library and a CLI. Both functionalities are
implemented as a part of \texttt{qbench} Python package. The exposed CLI tool is also named
\texttt{qbench}. For brevity, we do not discuss the exact structure of the package here, and instead limit ourselves to summarizing the architecture on the diagram in Fig. \ref{fig:architecture}. For further details, we refer an interested reader to the source code available at GitHub \cite{pyqbenchgithub} or at the
reference manual \cite{pyqbenchdocs}.

@@

PyQBench can be installed from official Python Package Index (PyPI) by running \texttt{pip install
pyqbench}. In a properly configured Python environment the installation process should also make the
\texttt{qbench} command available to the user without a need for further configuration.

@@

When using PyQBench as library, instances of Qiskit backends can be passed to functions that expect
them as parameters. However, in CLI mode, the user has to provide a YAML file describing the
backend. An example of such file can be found in Section \ref{sec:examples}, and the detailed
description of the expected format can be found at PyQBench's documentation.

@@

The Command Line Interface (CLI) of PyQBench has nested structure. The general form of the CLI
invocation is shown in listing \ref{lst:cli}.
\begin{lstlisting}[caption=Invocation of \texttt{qbench} script, label=lst:cli]
qbench <benchmark-type> <command> <parameters>
\end{lstlisting}
Currently, PyQBench's CLI supports only one type of benchmark (discrimination of parametrized
Fourier family of measurements), but we decided on hierarchically structuring the CL to allow for
future extensions. Thus, the only accepted value of \texttt{<benchmark-type>} is
\texttt{disc-fourier}.
%
The \texttt{qbench disc-fourier} command has four subcommands:

@@

\subsection{Using \texttt{qbench} CLI}
PyQBench offers a simplified way of conducting benchmarks using a Command Line Interface (CLI).
The workflow with PyQBench's CLI can be summarized as the following list of steps:

\begin{enumerate}
  \item Preparing configuration files describing the backend and the experiment scenario.
  \item Submitting/running experiments. Depending on the experiment scenario, execution can be
		synchronous, or asynchronous.
  \item (optional) Checking the status of the submitted jobs if the execution is asynchronous.
  \item Resolving asynchronous jobs into the actual measurement outcomes.
  \item Converting obtained measurement outcomes into tabulated form.
\end{enumerate}
\subsubsection{Preparing configuration files}
The configuration of PyQBench CLI is driven by YAML files. The first configuration file describes
the experiment scenario to be executed. The second file describes the backend. Typically, this
backend will correspond to the physical device to be benchmarked, but for testing purposes, one
might as well use any other Qiskit--compatible backend including simulators. Let us first describe
the experiment configuration file, which might look as follow.
\begin{lstlisting}[language=Python, caption=Defining the experiment, label=lst:experiment]
type: discrimination-fourier
qubits:
	- target: 0
	  ancilla: 1
	- target: 1
      ancilla: 2
angles:
	start: 0
	stop: 2 * pi
	num_steps: 3
gateset: ibmq
method: direct_sum
num_shots: 100
\end{lstlisting}

The second configuration file describes the backend. We decided to decouple the experiment and the
backend files because it facilitates their reuse. For instance, the same experiment file can be used
to run benchmarks on multiple backends, and the same backend description file can be used with
multiple experiments.

@@

\subsubsection{Running the experiment and collecting measurements data}
After preparing YAML files defining experiment and backend,
running the benchmark can be launched by using the following command line invocation:
\begin{lstlisting}[language=Python]
qbench disc-fourier benchmark experiment_file.yml backend_file.yml
\end{lstlisting}
The output file will be printed to stdout. Optionally, the \texttt{- -output OUTPUT} parameter might
be provided to write the output to the \texttt{OUTPUT} file instead.
\begin{lstlisting}[language=Python]
qbench disc-fourier benchmark experiment_file.yml backend_file.yml --output async_results.yml
\end{lstlisting}
The result of running the above command can be twofold:
\begin{itemize}
	\item If the backend is asynchronous, the output will contain intermediate data containing, amongst
		others, \texttt{job\_ids} correlated with the circuit they correspond to.
	\item If the backend is synchronous, the output will contain measurement outcomes (bitstrings)
		for each of the circuits run.
\end{itemize}

@@

\subsubsection{Resolving asynchronous jobs}
For asynchronous experiments, the stored intermediate data has to be resolved in actual
measurements' outcomes. The following command will wait until all jobs are completed and then write
a result file.
\begin{lstlisting}[language=Python]
qbench disc-fourier resolve async-results.yml resolved.yml
\end{lstlisting}
The resolved results, stored in \texttt{resolved.yml}, would look just like if the experiment was
run synchronously. Therefore, the final results will look the same no matter in which mode the
benchmark was run, and hence in both cases the final output file is suitable for being an input for
the command computing the discrimination probabilities.

@@

\section{Impact}
With the surge of availability of quantum computing architectures in recent years it becomes
increasingly difficult to keep track of their relative performance. To make this case even more
difficult, various providers give access to different figures of merit for their architectures. Our
package allows the user to test various architectures, available through \texttt{qiskit} and Amazon
BraKet using problems with simple operational interpretation. We provide one example built-in in the
package. Furthermore, we provide a powerful tool for the users to extend the range of available
problems in a way that suits their needs.

Due to this possibility of extension, the users are able to test specific aspects of their
architecture of interest. For example, if their problem is related to the amount of coherence (the
sum of absolute value of off-diagonal elements) of the states present during computation, they are
able to quickly prepare a custom experiment, launch it on desired architectures, gather the result,
based on which they can decide which specific architecture they should use.

>>>>

Noisy Intermediate-Scale Quantum (NISQ)~\cite{preskill} devices are storming the market, with a wide
selection of devices based on different architectures and accompanying software solutions. Among
hardware providers offering public access to their gate--based devices, one could mention Rigetti
\cite{rigetti}, IBM \cite{ibmq}, Oxford Quantum Group \cite{oxforf}, IonQ \cite{ionq} or Xanadu
\cite{xanadu}. Other vendors offer devices operating in different paradigms. e.g.  D-Wave \cite{dwave}
and their quantum annealers, or QuEra devices \cite{quera} based on neural atoms. Most vendors provide their own
software stack for interacting with their devices. Rigetti's computers are available through Forest
SDK \cite{sdk} and PyQuil library \cite{pyquil} and IBM Q \cite{ibmq} computers can be accessed
through Qiskit \cite{qiskit} or IBM Quantum Experience web interface \cite{ibmqplatform}. Amazon Braket
\cite{amazon} offers access to several quantum devices under a unified API. There are also several
libraries able to integrate with multiple hardware vendors. Examples include IBM Q's Qiskit, Zapata Computing's
Orquestra \cite{zapata}, XACC \cite{xacc} and NVIDIA's CUDA Quantum \cite{cudaquantum}.

It is well known that NISQ devices have their limitations \cite{preskillnew}. The question is to
what extent those devices can perform meaningful computations? To answer this question, one has to
devise a methodology for benchmarking them. For gate--based computers, on which this paper focuses,
there already exist several approaches. One could mention randomized benchmarking
\cite{liu2022sampling, knill2008randomized, wallman2014randomized, helsen2022general,
cornelissen2021scalable}, benchmarks based on the quantum volume \cite{cross2019validating,
moll2018quantum, pelofske2022volume} or quantum tomography \cite{bantysh2021quantum}.

In this paper, we introduce a different approach to benchmarking gate--based devices with a simple
operational interpretation. In our method, we test how good the given device is at guessing which of
the two known von Neumann measurements were performed during the experiment. We implemented our
approach in an open-source Python library PyQBench. The library supports any device available
through the Qiskit library, including all  IBM Q or Amazon Braket.
Along with the library, the PyQBench package contains a command line tool for running the most common
benchmarking scenarios.

@@

In this section, we describe how the benchmarking process in PyQBench works. We start by
discussing necessary mathematical preliminaries. Then, we present the general form of the
discrimination scheme used in PyQBench and discuss it in the context of current, limited NISQ devices.
We refer readers interested in more details concerning the mathematical foundations behind PyQBench to Section
\ref{app:preliminaries} in the supplementary material.

@@

A von Neumann measurement $\PP$ is a collection of rank--one projectors
$\{\proj{u_0}, \ldots, \proj{u_{d-1}}\}$ that sum up to the identity operator, i.e.
$ \, \, \sum_{i=0}^{d-1} \proj{u_i} = \1$. If $U$ is a unitary matrix of size $d$, one can construct
a von Neumann measurement $\PP_{U}$ by taking projectors onto its columns. In this case, we say that
$\PP_{U}$ is described by the matrix $U$.

@@

Without loss of generality\footnote{Explaining why we can consider only discrimination scheme
between $\PP_{\Id}$ and $\PP_{U}$ is beyond the scope of this paper. See
\cite{puchala2018strategies} for a in depth explanation.}, we consider discrimination task between
single qubit measurements $\PP_\Id$, performed in the Z-basis, and an alternative
measurement $\PP_U$ performed in the basis $U$. 
The scheme presented in Fig.~\ref{fig:theoretical_scheme} requires an
auxiliary qubit. First, the joint system is prepared in some state $\ket{\psi_0}$. Then, one of the
measurements, either $\PP_U$ or $\PP_\1$, is performed on the first part of the system. Based on its
outcome $i$, we choose another binary measurement $\mathcal{P}_{V_i}$ and perform it on the second qubit,
obtaining the outcome $j$. Finally, if $j=0$, we say that the performed measurement is
$\mathcal{P}_U$, otherwise we say that it was $\mathcal{P}_\Id$.

Note, however, that the discrimination scheme described above can work regardless of the dimensionality
of the unitary $U$. The main difference is the dimension of the auxiliary system, which in general
can be larger than two. This dimension depends on the Schmidt rank of the optimal input state. However, 
for most discrimination schemes, the  Schmidt rank equals at most two, and the auxiliary system
is also a qubit, see \cite{puchala2018strategies} for details. Note that the final measurement $\PP_{V_i}$ 
is always binary, independently of the dimension of the auxiliary system.

@@

Current NISQ devices are unable to perform conditional measurements, which is the biggest obstacle
to implementing our scheme on real hardware. We circumvent this problem by slightly
adjusting our scheme to only use components available on current devices. There are two possible options: using a postselection or a direct sum $V_0^\dagger\oplus V_1^\dagger$.

@@

	The first idea uses a postselection scheme. In the original scheme, we measure the first qubit
	and only then determine which measurement should be performed on the second one. Instead, we can
	run two circuits, one with $\PP_{V_0}$ and one with $\PP_{V_1}$ and measure both
	qubits. We can then discard the results of the circuit for which label $i$ does not match measurement
	label $k$. The circuit for postselection looks as depicted in Fig.
	\ref{fig:postselection}.

@@

	One can see why such a circuit is equivalent to the original discrimination scheme.
	If we rewrite the block-diagonal matrix $V_0^\dagger \oplus V_1^\dagger$ as follows:
	\begin{equation}
		\label{eq:directsum}
		V_0^\dagger \oplus V_1^\dagger = \proj{0}\otimes V_0^\dagger + \proj{1} \otimes V_1^\dagger,
	\end{equation}
	we can see that the direct sum in Eq. \eqref{eq:directsum} commutes with the measurement on the
	first qubit. Thanks to this, we can switch the order of operations to obtain the circuit from
	Fig. \ref{fig:directsum}. Now, depending on the outcome $i$, one of the summands in
	Eq. \eqref{eq:directsum} vanishes, and we end up performing the same operations as in
	the original scheme.

@@

Compared to other approaches, our approach allows for a very simple operational
interpretation of the scheme and its results. This is especially useful for
newcomers to the field, who may be put off by more complicated approaches.
Another benefit is, especially for advanced users, the ability to control
resources utilized during the benchmarking. We can consider here resources such
as entanglement or coherence. Finally, the figure of merit we wish to calculate and
to which we compare the results is fairly simple to obtain. The main downside of
our approach is the exponential number of circuits we need to consider.

@@

\begin{enumerate}
	\item Assembling circuits for both postselection and direct sum--based discrimination schemes.
	\item Executing the whole benchmarking scenario on specified backend (either real hardware or
	software simulator).
	\item Interpreting the obtained outputs in terms of discrimination probabilities.
\end{enumerate}
Note that the execution of circuits by PyQBench is optional. The user might opt-in
for fine-grained control over the execution of the circuits. For instance, suppose the user wants to
simulate the discrimination experiment on a noisy simulator. In such a case, they can define the
necessary components and assemble the circuits using PyQBench. The circuits can then be altered,
e.g. to add noise to particular gates, and then run using any Qiskit backend by the user. Finally,
PyQBench can be used to interpret the measurements to obtain discrimination probability.

The PyQBench library also contains a readily available implementation of all necessary components
needed to run discrimination experiments for parametrized Fourier family of measurements (see
Section \ref{app:optimal-probability} in supplemental material). However, if one only wishes to use
this particular family of measurements in their benchmarks, then using PyQBench as a command line tool
is a more straightforward way. The PyQBench's CLI does not require writing Python code and is driven
by YAML \cite{yaml} files describing the benchmark to be performed and the description of the backend on which the benchmark should be run. The same benchmark can be used with different backends and vice versa.

@@

\subsubsection{Overview of the software structure}
PyQBench can be used both as a library and a CLI. Both functionalities are
implemented as a part of \texttt{qbench} Python package. The exposed CLI tool is also named
\texttt{qbench}. For brevity, we limit ourselves here to summarizing the architecture on the diagram in Fig. \ref{fig:architecture}. For further details, we refer an interested reader to the source code available at GitHub \cite{pyqbenchgithub} or at the reference manual \cite{pyqbenchdocs}.

@@

PyQBench can be installed from the official Python Package Index (PyPI) by running \texttt{pip install
pyqbench}. In a properly configured Python environment, the installation process should also make the
\texttt{qbench} command available to the user without a need for further configuration.

@@

When using PyQBench as a library, instances of Qiskit backends can be passed to functions that expect
them as parameters. In CLI mode, the user has to provide a YAML file describing the
backend. An example of such a file can be found in Section \ref{sec:examples}, and the detailed
description of the expected format can be found in  PyQBench's documentation.

@@

The Command Line Interface (CLI) of PyQBench has a nested structure. The general form of the CLI
invocation is shown in listing \ref{lst:cli}.
\begin{lstlisting}[caption=Invocation of \texttt{qbench} script, label=lst:cli]
qbench <benchmark-type> <command> <parameters>
\end{lstlisting}
Currently, PyQBench's CLI supports only one type of benchmark (discrimination of parametrized
Fourier family of measurements), but we decided on hierarchically structuring the CL to allow for
future extensions. Thus, the only accepted value of \texttt{<benchmark-type>} is
\texttt{disc-fourier}.
%
The \texttt{qbench disc-fourier} command has four subcommands:

@@

\subsection{Using \texttt{qbench} CLI}
The workflow with PyQBench's CLI can be summarized as the following list of steps:

\begin{enumerate}
  \item Preparing configuration files describing the backend and the experiment scenario.
  \item Submitting/running experiments. Depending on the experiment scenario, execution can be
		synchronous, or asynchronous.
  \item (optional) Checking the status of the submitted jobs if the execution is asynchronous.
  \item Resolving asynchronous jobs into the actual measurement outcomes.
  \item Converting obtained measurement outcomes into tabulated form.
\end{enumerate}
\subsubsection{Preparing configuration files}
The configuration of PyQBench CLI is driven by YAML files. The first configuration file describes
the experiment scenari. The second file describes the backend. Typically, this
backend will correspond to the physical device to be benchmarked, but for testing purposes, one
might as well use any other Qiskit--compatible backend including simulators. The experiment configuration
file, which might look as follows.
\begin{lstlisting}[language=Python, caption=Defining the experiment, label=lst:experiment]
type: discrimination-fourier
qubits:
	- target: 0
	  ancilla: 1
	- target: 1
      ancilla: 2
angles:
	start: 0
	stop: 2 * pi
	num_steps: 3
gateset: ibmq
method: direct_sum
num_shots: 100
\end{lstlisting}

The second configuration file describes the backend. We decided to decouple the experiment and the
backend files because it facilitates their reuse. The same experiment file can be used
to run benchmarks on multiple backends, and the same backend description file can be used with
multiple experiments.

@@

\subsubsection{Running the experiment and collecting measurements data}
After preparing YAML files defining an experiment and backend,
running the benchmark can be launched by using the following command line invocation:
\begin{lstlisting}[language=Python]
qbench disc-fourier benchmark experiment_file.yml backend_file.yml
\end{lstlisting}
The output file will be printed to stdout. Optionally, the \texttt{- -output OUTPUT} parameter might
be provided to write the output to the \texttt{OUTPUT} file instead.
\begin{lstlisting}[language=Python]
qbench disc-fourier benchmark experiment_file.yml backend_file.yml --output async_results.yml
\end{lstlisting}
The result of running the above command can be twofold:
\begin{itemize}
	\item If the backend is asynchronous, the output will contain intermediate data containing, amongst
		others, \texttt{job\_ids} correlated with the circuit they correspond to.
	\item If the backend is synchronous, the output will contain measurement outcomes (bitstrings)
		for each of the circuits run.
\end{itemize}

@@

\subsubsection{Resolving asynchronous jobs}
For asynchronous experiments, the stored intermediate data has to be resolved in actual
measurements' outcomes. The following command will wait until all jobs are completed and then write
a result file.
\begin{lstlisting}[language=Python]
qbench disc-fourier resolve async-results.yml resolved.yml
\end{lstlisting}
The resolved results, stored in \texttt{resolved.yml}, would look just like if the experiment was
run synchronously. Therefore, the final results will look the same no matter in which mode the
benchmark was run, and hence in both cases, the final output file is suitable for being an input for
the command computing the discrimination probabilities.

@@

\section{Impact}
With the surge of availability of quantum computing architectures in recent years it becomes
increasingly difficult to keep track of their relative performance. Moreover, various providers give access to different figures of merit for their architectures. Our
package allows the user to test various architectures, available through \texttt{qiskit} and Amazon
BraKet using problems with simple operational interpretation. We provide one example built-in in the
package. Furthermore, we provide a powerful tool for the users to extend the range of available
problems in a way that suits their needs.

Due to this possibility of extension, the users can test specific aspects of their
architecture of interest. For example, if their problem is related to the amount of coherence (the
sum of absolute value of off-diagonal elements) of the states present during computation, they are
able to quickly prepare a custom experiment, launch it on desired architectures, gather the result,
based on which they can decide which specific architecture they should use.

### 2023-10-05 txt/supplemental.tex
Cleanup manuscript and supplemental file
Temporarily change supplemental to verify CI works
Revert changes in supplemental material
Move section 2 to the appendix + conciseness pass

\section{Mathematical preliminaries} \label{app:preliminaries}

Let $\mathcal{M}_{d_1,d_2}$ be the set of all matrices of dimension $d_1 \times d_2$ over
the field $\mathbb{C}$. For  simplicity, square matrices will be denoted by
$\mathcal{M}_d$.
By $\Omega_d$, we will denote the set of quantum states, that is
positive semidefinite operators having trace equal to one.
The subset of $\mathcal{M}_d$ consisting of unitary matrices will be denoted
by $\UU_d$, while its subgroup of diagonal unitary operators will be denoted by
$\DD \UU_d$.

@@

>>>>

\section{Existing benchmarking methodologies and software}
Unsurprisingly, PyQBench is not the only software package for benchmarking gate--based devices.
While we believe that our approach has significant benefits over other benchmarking techniques, for
completeness, in this section we discuss some of the currently available similar software.

Probably the simplest benchmarking method one could devise is simply running known algorithms and
comparing outputs with the expected ones. Analyzing the frequency of the correct outputs, or the
deviation between actual and expected outputs distribution provides then a metric of the performance
of a given device. Libraries such as Munich Quantum Toolkit (MQT) \cite{mqt2022, mqt-bench} or
SupermarQ \cite{supermarq, supermarkq-github} contain benchmarks leveraging multiple algorithms,
such as Shor's algorithm or Grover's algorithm.
% or  implementing e.g. GHZ state \cite{ghz}.
Despite being intuitive and easily interpretable, such benchmarks may have some problems. Most
importantly, they assess the usefulness of a quantum device only for a very particular algorithm,
and it might be hard to extrapolate their results to other algorithms and applications. For
instance, the inability of a device to consistently find factorizations using Shor's algorithms does
not tell anything about its usefulness in Variational Quantum Algorithm's.

Another possible approach to benchmarking quantum computers is randomized benchmarking. In this
approach, one samples circuits to be run from some predefined set of gates (e.g. from the Clifford
group) and tests how much the output distribution obtained from the device running these circuits
differs from the ideal one. It is also common to concatenate randomly chosen circuits with their
inverses (which should yield the identity circuit) and run those concatenated circuits on the
device. Libraries implementing this approach include Qiskit \cite{qiskit-randomized} or PyQuil
\cite{forest-benchmarking}. Another, equally popular, benchmarking method is
quantum tomography \cite{bantysh2021quantum}. Implementation of quantum tomography for
benchmarking NISQ devices can be found in \cite{tomography}.

In \cite{patel2020experimental}, the authors evaluated several IBM-Q machines using seven benchmarks
taking into account the errors and execution time.

QASMBench \cite{li2023qasmbench} is one of the first benchmark suites aiming at evaluating NISQ
devices using quantum applications from a broad range of domains, mainly using an approach based on
fidelity estimation. The benchmark presented in the paper compares the fidelity of execution among
the IBM-Q machines, the IonQ QPU and the Rigetti Aspen M-1 system.

Another quantity used for benchmarking NISQ devices is quantum volume. The quantum volume
characterizes the capacity of a device for solving computational problems. It takes into account
multiple factors like the number of qubits, connectivity and measurement errors. The Qiskit library
allows one to measure the quantum volume of a device by using its \texttt{qiskit.ignis.verification.quantum\_volume}. Other implementations of Quantum Volume can be found as well, see e.g.
\cite{volume-in-practice}.

We should also mention cross-entropy
benchmarking~\cite{boixo2018characterizing}, which was utlilized in validation
of the Sycamore-53 QPU supremacy experiments~\cite{arute2019quantum}. In this
approach, the quality of an algorithm implemented on the QPU is measured by
calculating the cross entropy of bit-strings actually sampled from the QPU,
compared to ideal bitstrings.

Finally, it is worth pointing out there is an ongoing effort towards standardization
of benchmarking of quantum computers. In \cite{standardization}, the authors present
plans for designing a benchmarking suite based on measuring a set of standardized
key performance indicators (KPIs).

\section{Mathematical preliminaries} \label{app:preliminaries}
Let $\mathcal{M}_{d_1,d_2}$ be the set of all matrices of dimension $d_1 \times d_2$ over
the field $\mathbb{C}$. For  simplicity, square matrices will be denoted by
$\mathcal{M}_d$.
By $\Omega_d$, we will denote the set of quantum states, that is
positive semidefinite operators having trace equal to one.
The subset of $\mathcal{M}_d$ consisting of unitary matrices will be denoted
by $\UU_d$, while its subgroup of diagonal unitary operators will be denoted by
$\DD \UU_d$.

@@

\bibliographystyle{unsrt}
\bibliography{references}
